Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2022-05-06 FDT2020/response/response.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\newcommand{\qed}{\hspace*{\fill}\mbox{$\Box$}}
\newcommand{\choos}[2] {\left( \begin{array}{c} #1\\#2
\end{array} \right)}

%\input{psfig}

\begin{document} \bibliographystyle{alpha}
\def\proofend{\hfill$\Box$\medskip}
\def\Proof{\noindent{\bf Proof:\ \ }}
\def\Sketch{\noindent{\bf Sketch:\ \ }}
\def\eps{\epsilon}
\newcommand{\arash}[1]{\noindent{\bf {\color{blue!60!black}{\sc Arash:}  #1}}}
\newcommand{\cindy}[1]{\noindent{\bf {\color{magenta!80!black}{\sc Cindy:}  #1}}}

\newcommand{\response}[1]{\noindent{{\color{blue!80!black}  #1}}}
\renewcommand{\arash}[1]{}
\renewcommand{\cindy}[1]{}
\title{Responses to reviewers comments}
\maketitle

\section{General Response}
\response{We thank the reviewers for their insightful and constructive comments. We sincerely apologize for the long delay in submitting our revision.  We thank our editor very much for the extension. Our junior author finished his training and took a job in industry. The more senior co-author at a national lab accepts her considerable share of responsibility for the delay as well.
}
%\arash{TODO: Thank the reviewers at length. Apologies for the delay in submission for the revisions, thanks the editor for the extensions.}
\section{Detailed Responses}

\subsection{Reviewer 1}
\subsubsection{Major Comments}
\response{We have addressed the issue in the proof of Theorem 3 as detailed below. We also made several major revision in our presentation especially with regards to the more technical parts of the paper.}
\subsubsection{Minor Comments}
\begin{itemize}
\item Introduction: Introduction jumps between describing methods and problems the methods are applied to. There is a subsection 1.2 Experiments with more problem descriptions. Please restructure the introduction so
that the flow is more logical.
\response{We agree that the introduction is a little bit lengthy and not direct. However, we wanted to make sure that the reader is familiar with our terminology before diving into the results presentation.

@@

That being said, we have made some changes in the introduction. We hope the new state of the introduction is satisfactory.}

\item p. 2, last paragraph: “Integrality gap” is defined with respect to an IP instance, instead of a family of IP instances (IP formulation) for a problem. This makes the paragraph rather awkward to read, because the statements are not correct as stated.
\response{We made various changes starting with the paragraph before Definition 1 through the discussion that follows. We compare our Definition 1 to other definitions of the integrality gap. Our definition is the integrality gap of a polyhedron. We explain why we use this definition. We also added more context about our definition of integrality gap to the abstract, so the reader is aware immediately that we are using a polyhedron-based definition.}

@@

\item p. 3, below equation (1): Why is $A$ defined as a matrix of rational numbers but $b$ a vector or real numbers?

\response{Fixed. Both $A$ and $b$ are comprised of rational numbers.}
\item p. 3: Is there a reason why you choose to not include the coefficients of the objective function in an instance?
\response{Added this sentence: The reason that we define integrality gap for a polyhedron as opposed to an instance of a problem is that our algorithms are independent of the cost function $c$. }

\item p. 4, Definition 1: Now the integrality gap is defined with respect to a matrix $A$ and vector $b$, taking the supremum (not maximum, please correct) over all coefficients $c$, that are nonnegative. This is inconsistent
with the previous statement, and also nonstandard as far as I know. Please add some remarks for the reader about this choice. Also, the importance of considering $c \geq 0$ only should be emphasized more in the text (it is of
vital importance when using Carr-Vempala).

\response{As stated above, the discussion following Definition 1. We added a remark that $c$ must be non-negative for Carr-Vempala to work.}

\item p. 4, equation (4): supremum instead of maximum
\response{Fixed.}

\item p. 5, below Proposition 2: I don’t see that Proposition 2 and Theorem 3 are equivalent as stated. (Also in light of the remark about one result having an algorithm, and the other not.)
\response{Fixed. Now we say they are similar statement. We also tried to word the statements in a way to draw the similarity between the two.}

@@

\response{Thank you for pointing this out. We corrected to say $\tilde{z}^i\in \mathcal{D}(S)$. We also added this observation at the beginning of section 2 that is cross-reference in the proofs: Any point $x\in [0,1]^n$ that is a convex combination of points in $\mathcal{D}(S)$, is also a convex combination of points in $\mathcal{D}(S) \cap [0,1]^n$}
\item p. 11, third paragraph: Same problem as previous bullet point.
\response{Fixed with referencing the Observation above.}
\item p. 13, top of page: Same problem.
\response{Fixed with referencing the Observation above.}
\item p. 24 and following, references: missing capitalization in a few entries
\response{We fixed many instances of missing capitalization.}

\end{itemize}

\subsection{Reviewer 2}

\subsubsection{Major Comments}
The computational results seem quite weak to me. The classic Feasibility Pump heuristic is known to produce solutions of fairly poor quality; it is designed primarily to produce feasible solutions to potentially difficult MIPs
quickly. On TAP and 2EC, the authors compare to problem-specific approximation algorithms that are not particularly designed to produce high quality solutions either.
\response{We respect and understand this reviewer's point of view. We have added some caveats to the start of the experimental section. Our experiments are intended to demonstrate using FDT for experimentally studying integrality gaps. We have explicitly stated the we are not recommending FDT as a replacement heuristic. We hope it will complement other general heuristics. %Our comparison of FDT to feasibility pump is mainly to showcase a use case of the algorithm and a direction that could be possibly explored. We are not suggesting FDT as a replacement for FP. We will add remarks that make this clear to the reader.

With respect to Christofides, we mistakenly wrote that we compared FDT to Christofides.  In fact, we compare FDT with BOMC (Best of Many Christofides) - this has been fixed in the abstract and the paper. As presented in Figure 3, we don't settle for the worst-case upper bound of 3/2. Similarly, for TAP, FJ81 is the only rounding algorithm known for the solution of the CUT-LP. Again, we don't use the worst-case upper bound of 2. Instead we use the instance-specific gap provided by the algorithm. Also, FDT is a generic algorithm that applies to any formulation with a suitablly bounded integrality gap, while BOMC and FJ81 problem-specific.}

@@

\begin{enumerate}

\item The abstract is misleading in multiple ways. First, it says that the paper will give a constructive version of the Carr-Vempala theorem but does not mention the rather large loss in quality of $g$ to (potentially) $g^n$ . Secondly,
it says that “[Dom2IP] more quickly determines if an instance has an unbounded integrality gap” but really this algorithm will only sometimes provide proof that the integrality gap is unbounded; it is not guaranteed
to certify this. Lastly, it is claimed that Christofides’ algorithm is the “best previous approximation algorithm”. But Christofides is only best wrt. its theoretical approximation guarantee not its practical performance
which is what is compared in the paper.

\response{We made several edits to the abstract. We believe they address all the points mentioned above.}
\item The paper should mention the randomized metarounding algorithm by Carr and Vempala and compare it to the FDT algorithm. 
\response{We added this paragraph right before section 1.1: 
Carr and Vempala [CV00] showed if there is an algorithm $\mathcal{A}$ that for $I\in \mathcal{I}$ and $c\geq 0$ finds $z\in S(I)$ with $cz\leq g'cx$ for any $x\in P(I)$, then $g'x$ can be written as a convex combination of points in $S(I)$. However, for many known problems there is a gap between the best known $g'$ and $g(\mathcal{I})$. }
\item In the proofs of Lemma 8 and Lemma 11, I do not see why it can be assumed that $\lambda_0,\lambda_1,\lambda_2>0$. I think one has to allow the case in which
the branching step returns just 1 or (in the case of 2EC) just 2 vectors.
\response{Yes. We corrected this in the statement of Lemma 9, 10, 12. Also in the description of the Algorithm 2.}
\item As noted before, I find the comparisons against the FP heuristic on VC
and the 2-approximation for TAP rather dubious. To show that the FDT
algorithm performs much better than the theoretical $g^{|supp(x)|}$ bound, just comparing to the initial fractional solution is all that is needed. On the other hand, if the goal is to show that the FDT algorithm works well as a
general purpose feasibility heuristic, then comparing to a FDT to a more modern variant of FP (e.g. as in the Feasibility Pump 2.0 paper) would be more reasonable.

\response{We felt that readers might be curious about how FDT compares to another general incumbent heuristic. We were not able to find any code for feasibility pump 2.0. In the FP 2.0 paper, the authors wrote that the “overhead of constraint propagation can be quite large … Hence we need a very efficient implementation of the whole constraint propagation system, including both the single propagators and the overall propagation engine.”  FP 2.0 is much more complex than simple FP. To be fair to FP 2.0 in a comparison, we would need to program the authors’ optimizations to achieve their performance.  Since our goal was to show that the performance can be reasonable, we felt that implementing the full, optimized FP 2.0 was out of scope.}
%The comparisons with FP are not presented to claim that FDT to provide direct comparisons between the two algorithms, but to provide a baseline on the IG upper bounds provided by another LP-based algorithm. We want to emphasize that the scope of this work in mostly on the theoretical contributions, rather than experimental results.

@@

\end{enumerate}

\subsubsection{Minor Comments}
\begin{enumerate}

\item  The abstract says “Dom2IP” but the paper mostly uses “DomToIP”. 
\response{Fixed.}
\item In Theorem 3, I think $\mathcal{D}(S(I))$ should just be $S(I)$.
\response{The proof of this theorem only holds for the dominant of $S(I)$ and not $S(I)$ itself, although for many problems of interest we have $S(I)=\mathcal{D}(S(I))$. }

@@

\item On page 6, “which runs in polynomial time algorithm” should be “which runs in polynomial time”. 
\response{Fixed.}
\item On page 6, the use of $\theta_0$ and $\theta_1$ is not consistent with how $\gamma_0$ and $\gamma_1$ are
defined in Lemma 8. 
\response{Fixed and switched to $\gamma_0$ and $\gamma_1$ in the overview to be consistent with Lemma 8.}
\item In Theorem 6, is $E_x=\mbox{supp}(x)$? \response{Fixed. We added the definition of $E_x$ right before Theorem 6.}

\item On page 8, the wording “We use the idea of fundamental extreme point to create the hardest LP solutions to decompose.” seems to suggest that this is something that will be done in this paper.
\response{We changed the wording as follows: Known polyhedral structure makes it easier to study integrality gaps for such problems. In particular, for the subtour-elimination relaxation [CR98, BC11, CV04] introduced fundamental extreme points that create the ``hardest'' LP solutions to decompose.}.
\item On page 8, you refer to Figure 2 but this figure only shows up on page 23.

\response{The figure is moved to page 3.}
\item You mention that the Dom2IP algorithm can be used to prove that an IPs integrality gap is unbounded. How effective is this?

\response{
DomToIP starts from a feasible solution to the LP relaxation and tries to find an integer feasible solution in the support of the fractional solution. If Dom2IP fails to do this for any instance, we can conclude the integrality gap is infinite. We could perform a broad search over possible (useful/interesting) feasible points to the LP relaxation to find such a certificate for the unboundedness of the integrality gap.}
\item Lemma 9 requires $\gamma_0 + \gamma_1 \leq 1$. Perhaps this should already be included
in Lemma 8.
\response{Fixed. Also added a constraint into the formulation of LPC that ensures $\lambda_0 + \lambda_1\leq 1$ and corresponding changes in the proof of the Claim.}
\item In the proof of Lemma 9, “We need to show that $\gamma_0\hat{x}^0_j + \gamma\hat{x}^1_j \leq gx'_j.$" has
an extra $g$ in it.
\response{Fixed.}
\item On page 13, “nodes in level $L_i$” should be “nodes in level $i$” and “points
in the leaves” should be just “leaves.” \response{Fixed.}
\item On page 24, “applying this tool . . . can be a tool” is redundant. \response{Fixed.}
\end{enumerate}

@@

\response{Here the goal is to establish that there is a $z\leq x^{(\ell+1)}$. For the coordinates of $x^{(\ell+1)}$ that are zero we show the proof. The other coordinates of $x^{(\ell+1)}$ are 1. So we just need to emphasize that all the $\tilde{z}_i \leq 1$. This is added now, an a reference to an observation we added (Observation 7)}.

\item (12,10) “maintainS”  \response{Fixed.}
\item (13,28) “Growing and Pruning THE FDT tree”
\response{Fixed.}
\item (14,12) “It is easy to check” <– you do this quite a bit. Please go through your paper and
pay attention to all the “easy to check”, “obvious”, “clearly” etc. If these things really are easy, then just state them and don’t waste space. However, if some observation is necessary, think about the clearest and most concise way to impart this observation to the reader, and then do that instead of telling the reader that it should be “easy”.

\response{Thank you for the reminder. We find several occurrences of trivially, obviously, easy to check and clearly.}

\item (14,12) “THE set $L'$” \response{Fixed.}
\item (14,25) “From THE leaves” 
\response{Fixed.}
\item (16,11-17) You should discuss why the intuitive approach of binarizing an LP (just split each variable into two binary variables and copy the respective column of the
constraint matrix) does not lead to the same or better result. Or maybe it does?
\response{We added this paragraph in the discussion of why $\{0,1,2\}$ problems are different from binary problems: Another approach to a $\{0,1,2\}$ problem could be to ``binarize'' the variables, but splitting each variable into two binary copies and solving the binary problem via FDT for binary IPs (Theorem \ref{FDTFull}). One potential issue that can arise by doing this is that the binary version of a problem might have a larger integrality gap compared to the one that allows larger multiplicity. One example of this is the 2EC problem considered in this section. As mentioned above, we have $g(\mbox{2EC})\leq \frac{3}{2}$, while $g(\mbox{2ECSS})\geq \frac{3}{2}$, where 2ECSS is a problem equivalent to 2EC but without any doubled edges allowed {CKKK08}.}

### 2022-05-06 FDT2020/Introduction.tex
added a word
final commit for revisions

%\arash{I added this figure \ref{fig:CVpoint} but maybe we can put it later in the experiments section to make the introduction shorter?} \cindy{Yes that would be fine, if the reader could also use that figure then.  They could page forward to see the figure when reading the introduction.}
%\arash{I changed the paragraph below a little bit, can you check?} \cindy{Looks good. I just changed one sentence to the active voice.  If you're OK with that, you can delete these 2 comments. Thanks!}
Carr and Vempala \cite{Carr2004} showed that $g(\2ec)$ is achieved for instances where the optimal solution to $\min_{x\in \subtour(G)}cx$ is a Carr-Vempala point. Multiple groups of researchers have conjectured that $g(\2ec)\leq \frac{6}{5}$ (see \cite{alexander2006integrality,boydlegault,hn19}, but the only known upper bound on $g(\2ec)$ stands at $\frac{3}{2}$ \cite{wolsey}. We show that the integrality gap is at most $\frac{6}{5}$ for Carr-Vempala points with at most 12 vertices on the cycle formed by the fractional edges, thereby confirming the conjecture for these instances. Note that the number of vertices in these instances can be arbitrarily high since the paths of edges with $x$-value 1 can be arbitrarily long.
\cindy{This sounds like a contribution of the paper.} \arash{Yes, particlarly since a similar research question have been studied by \cite{alexander2006integrality} and \cite{TSPcompute}}
\cindy{We probably need an explicit list of contributions.  It could be another subsection of the introduction.  This will help the reviewer and eventually, if published, the reader.}
\subsection{Contribution summary}
In summary, this papers's contributions are:
\begin{itemize}
\item We give an algorithm to express any feasible point for the LP relaxation of a binary IP, when scaled by a factor depending on the integrality gap of the IP formulation, 
%(or a \{0,1,2\} IP) 
as a convex combination of feasible solutions, provided the IP integrality gap is bounded.  This is the Fractional Decomposition Tree (FDT) algorithm.
\item We give a simple  algorithm to give a feasible solution for any binary IP with bounded integrality gap. If this algorithm fails, then the integrality gap is infinite.
\item We experimentally show that FDT can give good approximate solutions to small instances of vertex-cover and tree augmentation problems.
\item We show that the integrality gap for 2EC (minimum 2-edge-connected sub-multigraph) is at most $\frac{6}{5}$ for Carr-Vempala points with at most $12$ vertices on the cycle formed by the fractional edges, matching the lower bound for the integrality gap for 2EC for these challenging instances.
\item We demonstrate (with 2EC) how FDT experiments can support theoretical work on integrality gaps for specific problems. In particular, we computationally prove an upper bound on small instances of 2EC with a specific structure.
\item We demonstrate (with VC) how FDT can be used to find feasible solutions of improved quality compared to those from the feasiblity pump \cite{fp1}.
\end{itemize}

>>>>

%\arash{I added this figure \ref{fig:CVpoint} but maybe we can put it later in the experiments section to make the introduction shorter?} \cindy{Yes that would be fine, if the reader could also use that figure then.  They could page forward to see the figure when reading the introduction.}
%\arash{I changed the paragraph below a little bit, can you check?} \cindy{Looks good. I just changed one sentence to the active voice.  If you're OK with that, you can delete these 2 comments. Thanks!}
Carr and Vempala \cite{Carr2004} showed that $g(\2ec)$ is achieved for instances where the optimal solution to $\min_{x\in \subtour(G)}cx$ is a Carr-Vempala point. Multiple groups of researchers have conjectured that $g(\2ec)\leq \frac{6}{5}$ (see \cite{alexander2006integrality,boydlegault,hn19}, but the only known upper bound on $g(\2ec)$ stands at $\frac{3}{2}$ \cite{wolsey}. We show that the integrality gap is at most $\frac{6}{5}$ for Carr-Vempala points with at most 12 vertices on the cycle formed by the fractional edges, thereby confirming the conjecture for these instances. Note that the number of vertices in these instances can be arbitrarily high since the paths of edges with $x$-value 1 can be arbitrarily long.
\cindy{This sounds like a contribution of the paper.} \arash{Yes, particlarly since a similar research question have been studied by \cite{alexander2006integrality} and \cite{TSPcompute}}
\cindy{We probably need an explicit list of contributions.  It could be another subsection of the introduction.  This will help the reviewer and eventually, if published, the reader.}
\subsection{Contribution summary}
In summary, this papers's contributions are:
\begin{itemize}
\item We give an algorithm to express any feasible point for the LP relaxation of a binary IP, when scaled by a factor depending on the integrality gap of the IP formulation, 
%(or a \{0,1,2\} IP) 
as a convex combination of feasible solutions, provided the IP integrality gap is bounded.  This is the Fractional Decomposition Tree (FDT) algorithm.
\item We give a simple  algorithm to give a feasible solution for any binary IP with bounded integrality gap. If this algorithm fails, then the integrality gap is infinite.
\item We experimentally show that FDT can give good approximate solutions to small instances of vertex-cover (VC) and tree augmentation problems.
\item We show that the integrality gap for 2EC (minimum 2-edge-connected sub-multigraph) is at most $\frac{6}{5}$ for Carr-Vempala points with at most $12$ vertices on the cycle formed by the fractional edges, matching the lower bound for the integrality gap for 2EC for these challenging instances.
\item We demonstrate (with 2EC) how FDT experiments can support theoretical work on integrality gaps for specific problems. In particular, we computationally prove an upper bound on small instances of 2EC with a specific structure.
\item We give proof-of-concept experimental results that demonstrate that FDT might be a reasonable heuristic in practice for some problems. For VC, FDT gives better solutions than the original feasibility pump \cite{fp1}, though requiring more time.  Similarly, for TAP, FDT usually gives much better solutions than the known 2-approximation \cite{FJ81}, but runs more slowly.

\end{itemize}
